### 2018-01-18 includes/abstract.tex
Applying paper feedback.

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and practically useful. Moreover a standard input language, SMT-LIB, has been developed, and many mature solvers support it. In this paper we present StringFuzz: a fuzzer and SMT-LIB problem generator for string solvers. With it we analyse four mature string solvers: Z3str3, Norn, CVC4, and ABC. We use it to craft inputs that elicit poor performance in Z3str3 and CVC4, and provide algorithmic analyses of the causes of the performance degradations. Finally, we present several minor bugs in Z3str3 that were exposed by StringFuzz, to justify its further use by other solver authors to likewise improve their solvers.

>>>>

    In this paper we present \fuzzer{}: an \smtlib{} problem fuzzer and generator for string solvers. We provide a rich repository of problems we generated with \fuzzer{} in \smt{} format. We also analyse \numSolvers{} mature string solvers: \theSolvers{}. We use \fuzzer{} to craft inputs that elicit poor performance in \us{} and \cvc{}, and analyse the causes of the performance degradations. We then describe an algorithm to mitigate the performance degradation in \us{}. Finally, we describe several bugs in \us{} that were exposed by \fuzzer{}.

### 2018-01-18 includes/experiment.tex
Applying paper feedback.

    Here we present two particular sets of problems generated by \fuzzer{}: Eager-Hard and Lazy-Hard. These two problem sets were unique in their ability to elicit poor performance in one solver, but not another. Full details of test scripts, configurations, and problems that were used in the experiment can be found in the Appendix.

@@

        \subsubsection{Performance on Lazy-Hard}

            On this problem set \cvc{} did poorly, but \us{} did well. The graphs in Figures~\ref{fig:z3str3-lazy-hard} and ~\ref{fig:cvc4-lazy-hard} show the performance results for both solvers. We produce and analyse the execution trace files both solvers' execution on the problem set.

            \begin{figure}[h]
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/z3str3-lazy-hard.png}
                    \caption{Performance of \us{} on Lazy-Hard}
                    \label{fig:z3str3-lazy-hard}
                \end{subfigure}
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/cvc4-lazy-hard.png}
                    \caption{Performance of \cvc{} on Lazy-Hard}
                    \label{fig:cvc4-lazy-hard}
                \end{subfigure}
            \end{figure}

@@

        \subsubsection{Performance on Eager-Hard}

            On this problem set \cvc{} did well, but \us{} did poorly. The graphs in Figures~\ref{fig:z3str3-eager-hard} and ~\ref{fig:cvc4-eager-hard} show the performance results for both solvers. We produce and analyse the execution trace files both solvers' execution on the problem set, and present them in the Appendix.

            \begin{figure}[h]
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/z3str3-eager-hard.png}
                    \caption{Performance of \us{} on Eager-Hard}
                    \label{fig:z3str3-eager-hard}
                \end{subfigure}
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/cvc4-eager-hard.png}
                    \caption{Performance of \cvc{} on Eager-Hard}
                    \label{fig:cvc4-eager-hard}
                \end{subfigure}
            \end{figure}

@@

        \subsubsection{Concats-Extracts}

            No data on this yet.

@@

        This section presents our analysis of the observed behaviour of \cvc{} and \us{} on Lazy-Hard and Eager-Hard problems.

        On Lazy-Hard, \cvc{} does poorly and \us{} does well because ...

        On Eager-Hard, \cvc{} does well and \us{} does poorly because ...

>>>>

    Here we present two particular sets of problems generated by \fuzzer{}: \cHard{} and \zHard{}. These two problem sets were unique in their ability to elicit poor performance in one solver, but not another. Full details of test scripts, configurations, and problems that were used in the experiment can be found in the Appendix.

@@

        \subsubsection{Performance on \cHard{}}

            On this problem set \cvc{} did poorly, but \us{} did well. The graphs in Figures~\ref{fig:z3str3-cvc4-hard} and ~\ref{fig:cvc4-cvc4-hard} show the performance results for both solvers. We produce and analyse the execution trace files both solvers' execution on the problem set.

            \begin{figure}[h]
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/cvc4-cvc4-hard.png}
                    \caption{Performance of \cvc{} on \cHard{}}
                    \label{fig:cvc4-cvc4-hard}
                \end{subfigure}
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/z3str3-cvc4-hard.png}
                    \caption{Performance of \us{} on \cHard{}}
                    \label{fig:z3str3-cvc4-hard}
                \end{subfigure}
                \caption{Results on \cHard}
                \label{fig:cvc4-hard}
            \end{figure}

@@

        \subsubsection{Performance on \zHard{}}

            On this problem set \cvc{} did well, but \us{} did poorly. The graphs in Figures~\ref{fig:z3str3-z3str3-hard} and ~\ref{fig:cvc4-z3str3-hard} show the performance results for both solvers. We produce and analyse the execution trace files both solvers' execution on the problem set, and present them in the Appendix.

            \begin{figure}[h]
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/cvc4-z3str3-hard.png}
                    \caption{Performance of \cvc{} on \zHard{}}
                    \label{fig:cvc4-z3str3-hard}
                \end{subfigure}
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/z3str3-z3str3-hard.png}
                    \caption{Performance of \us{} on \zHard{}}
                    \label{fig:z3str3-z3str3-hard}
                \end{subfigure}
                \caption{Results on \zHard}
                \label{fig:z3str3-hard}
            \end{figure}

@@

        \subsubsection{Performance Concats-Extracts}

            \todo{}: No data on this yet.

@@

        This section presents our analysis of the observed behaviour of \cvc{} and \us{} on \cHard{} and \zHard{} problems.

        \todo{}: On \cHard{}, \cvc{} does poorly and \us{} does well because ...

        \todo{}: On \zHard{}, \cvc{} does well and \us{} does poorly because ...

### 2018-01-18 includes/introduction.tex
Applying paper feedback.

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and practically useful. Moreover a standard input language, \smt{}, has been developed and many mature solvers support it.

    In this paper, we present the following contributions:

    \begin{enumerate}
        \item \fuzzer{}: a fuzzer of \smt{} string and regex problems.
        \item Generated inputs that elicit poor performance in \us{} and \cvc{}.
        \item A recommended fix to \us{} that alleviates the poor performance.
        \item A large suite of \smt{} string and regex problems we generated with \fuzzer{}.
    \end{enumerate}

>>>>

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and practically useful. Moreover a standard input language, \smt{}\cite{smt}, has been developed and many mature solvers support it.

    Our contributions follow. We describe the \fuzzer{} tool: a fuzzer of \smt{} string and regex problems. We then present a large suite of \smt{} string and regex problems we generated with \fuzzer{}. Then we describe two specific inputs - \cHard{} and \zHard{} - and experiments that show uniquely poor performance in \cvc{} and \us{} on those inputs, respectively. Finally, we recommend an algorithm to alleviate the poor performance in \us{}.

### 2018-01-18 includes/prior.tex
Applying paper feedback.

    There are several fuzzers and problem generators currently available, but none of them support string or regex generation or transformation. For example the FuzzSMT [ref] tool generates \smt{} problems with bit vectors and arrays, but does not support strings or regexes. The SMTpp [ref] tool preprocesses and simplifies problems, but does not generate new ones or fuzz existing ones.

@@

>>>>

    There are several fuzzers and problem generators currently available, but none of them support string or regex generation or transformation. For example the FuzzSMT\cite{fuzzsmt} tool generates \smt{} problems with bit vectors and arrays, but does not support strings or regexes. The SMTpp\cite{smtpp} tool preprocesses and simplifies problems, but does not generate new ones or fuzz existing ones.

@@

    \hfill \\
    \todo{}: references

### 2018-01-18 includes/problems.tex
Applying paper feedback.

    \subsection{\fuzzer{}-Generated}

        This suite contains the following problem sets:

        \begin{itemize}
            \item Lengths
            \item Lengths-Concats
            \item Concats
            \item Concats-Extracts
            \item Overlaps
            \item Regex
            \item Many-Regexes
            \item Regex-Deep
            \item Regex-Pair
            \item Regex-Lengths
            \item Random??
        \end{itemize}

@@

\begin{verbatim}
while true; do stringfuzzg --random regex | z3str3 -in; done
\end{verbatim}

    \subsection{\fuzzer{}-Transformed}

        This suite contains the following problem sets:

        \begin{itemize}
            \item Amazon??
            \item Regex??
        \end{itemize}

@@

        \begin{enumerate}
            \item length lookup using eqc root
            \item length-testing bug
        \end{enumerate}

>>>>

    \subsubsection{\fuzzer{}-Generated}

        This suite contains the following problem sets: \\

        \begin{tabular}{|c|c|}
            \hline
            \textbf{Name}    & \textbf{Description} \\ \hline
            Lengths          & \todo{} \\ \hline
            Lengths-Concats  & \todo{} \\ \hline
            Concats          & \todo{} \\ \hline
            Concats-Extracts & \todo{} \\ \hline
            Overlaps         & \todo{} \\ \hline
            Regex            & \todo{} \\ \hline
            Many-Regexes     & \todo{} \\ \hline
            Regex-Deep       & \todo{} \\ \hline
            Regex-Pair       & \todo{} \\ \hline
            Regex-Lengths    & \todo{} \\ \hline
            Random??         & \todo{} \\ \hline
        \end{tabular}

        \hfill \\

@@

        {\scriptsize\begin{verbatim}
while true; do stringfuzzg --random regex | z3str3 -in; done\end{verbatim}}

        Some simple UNIX scripting can yield more interesting results. For example the following script finds a regex problem on which \us{} times out after 5 seconds:

        {\scriptsize\begin{verbatim}while true; do
    stringfuzzg --random regex --coincide --num-regexes 2 --num-terms 5 --depth 3 > hard-problem.smt2
    result=`z3str3 -T:5 hard-problem.smt2`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat hard-problem.smt2
        break
    fi
done\end{verbatim}}

    \subsubsection{\fuzzer{}-Transformed}

        This suite contains the following problem sets: \\

        \begin{tabular}{|c|c|}
            \hline
            \textbf{Name} & \textbf{Description} \\ \hline
            Amazon??      & \todo{} \\ \hline
            Regex??       & \todo{} \\ \hline
        \end{tabular}

        \hfill \\

@@

        \todo{}: Describe the length lookup bug and length-testing bug, giving the before/after commits.

### 2018-01-18 includes/stringfuzz.tex
Applying paper feedback.

            \begin{tabular}{|c|c|}
                \hline
                \textbf{Name} & \textbf{Description} \\ \hline
                concats       & \\ \hline
                lengths       & \\ \hline
                overlaps      & \\ \hline
                regex         & \\ \hline
                random-text   & \\ \hline
            \end{tabular}

@@

            \begin{tabular}{|c|c|}
                \hline
                \textbf{Name} & \textbf{Description} \\ \hline
                fuzz          & \\ \hline
                graft         & \\ \hline
                multiply      & \\ \hline
                nop           & \\ \hline
                reverse       & \\ \hline
                rotate        & \\ \hline
                translate     & \\ \hline
                unprintable   & \\ \hline
            \end{tabular}

@@

        \item[\texttt{stringmerge}] \hfill \\
            This tool takes as input two \smt{} problems and combines them into one problem.

>>>>

            \begin{tabular}{|c|c|}
                \hline
                \textbf{Name} & \textbf{Description} \\ \hline
                concats       & \todo{} \\ \hline
                lengths       & \todo{} \\ \hline
                overlaps      & \todo{} \\ \hline
                regex         & \todo{} \\ \hline
                random-text   & \todo{} \\ \hline
            \end{tabular}

@@

            \begin{tabular}{|c|c|}
                \hline
                \textbf{Name} & \textbf{Description} \\ \hline
                fuzz          & \todo{} \\ \hline
                graft         & \todo{} \\ \hline
                multiply      & \todo{} \\ \hline
                nop           & \todo{} \\ \hline
                reverse       & \todo{} \\ \hline
                rotate        & \todo{} \\ \hline
                translate     & \todo{} \\ \hline
                unprintable   & \todo{} \\ \hline
            \end{tabular}

@@

        \item[\texttt{stringmerge}] \hfill \\
            This tool takes as input two \smt{} problems and combines them into one problem. \textbf{TODO: Add details about merging strategies.}
